---
title:  "BUSINFO 704 2025 Q3 Week 7 Classification of  Credit Card Defaults"
author: "Dr Sarah Marshall"
format: 
  html:
    toc: true
    toc-depth: 3
    number-sections: true
    number-depth: 3
    toc-location: right
execute:
  cache: false  
editor: 
  markdown: 
    wrap: 72
---

# Overview {.unnumbered}

This document demonstrates the parameter tuning for three algorithms: Neural Networks, Support Vector Machines and LightGBM using credit card default data. 


The original version of this file uses a sample of 1000 observations (from the available 10000). It also uses under-sampling (`step_downsample`). These decisions speed up the computation time, however also reduce model accuracy. Alternative modelling decisions will be explored in the lab.
  
Note: this document is designed to document the workflow, rather than
communicate results in a report format. As such, code and output will be
shown in the rendered document.

```{r}
#| label: setup
#| echo: true
#| results: hide
#| message: false

# Set options

name_of_this_file <- "week07_example_default"  
figprefix = paste('Rfigs/', paste0(name_of_this_file, "_"), sep="")


# set options for installing packages 
# TRUE: packages will get downloaded during knitting (not recommended)
# FALSE: run code below to check required packages are installed prior to knitting.
install_packages <- FALSE  

# set options for rendering
# more info: https://yihui.org/knitr/options/#chunk-options
knitr::opts_chunk$set(
               echo = TRUE,        # show code
               message = FALSE,    # hide messages
               warning = FALSE,    # hide warnings
               fig.path=figprefix, # save figures
               fig.align='center', # figure alignment
               fig.show='hold',    # print figures at the end of each chunk
               out.width="70%",    # figure width, % of page
               fig.width=8,        # figure width
               fig.height=4.5,     # figure height
               size='footnotesize' # Font size for chunk output
               )
               
# set options for printing within rendered document
opt_width = 70 #60
options(width = opt_width,  
        pillar.print_min = 6,  
        pillar.print_max = 10
        )


# Required packages
packages_needed <-  c( 
        #"tidyverse",
        "GGally",
        "tidymodels",        # modelling functions
        "ISLR",              # data
        "themis",            # recipe steps for unbalanced data
        "NeuralNetTools",    # Neural networks
        "nnet",
        "vip",
        "lightgbm", "bonsai" # lightgbm        
        )

# Identify required packages that are not currently installed
packages_to_install <- packages_needed[
                          !packages_needed %in% installed.packages()
                          ]

# Install required packages
if(install_packages){
  sapply(packages_to_install, 
         install.packages,
         dependencies=TRUE, 
         repos="https://cloud.r-project.org")
}

# load all required packages
sapply(packages_needed, require, character=TRUE)

# request that tidymodels functions are given preference
tidymodels_prefer()

```



# Data 


## Data Preparation

 

```{r}
#| label: load-data
library(ISLR)                # 加载包
set.seed(672025)

data(Default)

glimpse(Default)
```

## Data Understanding

```{r}
#| label: plot-data
#| out-width: "100%"

Default |> 
  ggpairs(aes(colour=default, alpha = 0.5))
```


There are two classes. Notice the class imbalance.

```{r}
Default |>  count(default)
```

We will re-order the factor levels for simplicity. 
We will also take a sample of the data to improve the initial run-time. 
It is recommended to run on the full dataset later.


```{r}
Default2 <- Default |> 
  mutate(default = factor(default, levels = c("Yes", "No"))) |> 
  slice_sample(n = 1000)
```

```{r}
Default2 |>  count(default)
```

# Pre-processing

Partition the data using 80% training, 20% testing.

Stratified based on class.

```{r}
#| label: data_partition

set.seed(234234)

# Split 80% of the data into the training set 
split <- initial_split(Default2, strata = default)


# Create data frames for the three sets:
train_data <- training(split)
test_data  <- testing(split)

# 5-fold CV
folds <- vfold_cv(train_data, v = 5, strata = default)
```



# Modelling 


## Recipe

```{r}
#| label: specify-recipe
install.packages("themis")  # 如果没装过
library(themis)             # 加载后才能用 step_downsample

default_rec <- recipe(default ~ ., data = train_data) |>
  #step_upsample(default, over_ratio = 1) |>   # note: increases computation time
  step_downsample(default, under_ratio = 1) |> # note: decreases amount of data used
  step_normalize(all_numeric_predictors()) |>    
  # create dummy variables for nominal predictors
  step_dummy(all_nominal_predictors())|> 
  # remove zero variance predictors
  step_zv(all_predictors())   


```


```{r}
default_prepped <- 
default_rec |> 
  prep() |> 
  bake(new_data = NULL) 

default_prepped |> glimpse()
default_prepped |>  count(default)
default_prepped  |>  summary()
```

Notice that the mean of the numeric variables is 0.


## Create model and workflow

### Neural Network


We will tune on 2 parameters: `hidden_units` and `penalty`.


* `hidden_units` represents the number of neurons in the hidden layer.  Higher values can capture more complex relationships, but can lead to over-fitting. Lower values result in a simpler model, but risk under-fitting.
* `penalty` represents the weight decay Higher values lead to a simpler model (more weight decay) and lower values provide more flexibility, but can lead to over-fitting. 


```{r}
#| label: nn_recipe_model

nn_spec <- 
  mlp(hidden_units = tune(), penalty = tune(), epochs = 500) |>
  set_engine("nnet", trace = FALSE) |>
  set_mode("classification")

nn_wf <- workflow() |>
  add_model(nn_spec) |>
  add_recipe(default_rec)

nn_grid <- grid_regular(hidden_units(), penalty(), levels = c(4, 3))



```


### SVM

We will tune on 2 parameters: `cost` and `rbf_sigma`.


* Higher values of `cost` mean a greater penalty for misclassification. If cost is too high, it can lead to over-fitting. Lower values of `cost` result in a simpler model, but risk underfitting.
* Higher values of `rbf_sigma` mean that a smaller radius is used, and the decision boundary is sensitive to individual observations. Can lead to over-fitting. Lower values provide a smoother decision boundary, but can lead to under-fitting.


```{r}
svm_spec <- 
  svm_rbf(cost = tune(), rbf_sigma = tune()) |>
  set_engine("kernlab") |>
  set_mode("classification")

svm_wf <- workflow() |>
  add_model(svm_spec) |>
  add_recipe(default_rec)

svm_grid <- grid_regular(cost(), rbf_sigma(), levels = 3)

```



### LightGBM

We will tune on 2 parameters: `learn_rate` and `trees`.

* `learn_rate` controls how much each tree contributes to the final model. Smaller values of `learn_rate` mean learning happens more slowly, but does provide better generalisation. Higher values mean faster learning, but can lead to over fitting
* `trees` controls the number of trees.


 
```{r}
#| label: lgbm-model-defn

lgbm_spec <- 
  boost_tree(
  learn_rate = tune(),
  trees = tune()
  ) |>
  set_engine("lightgbm") |>
  set_mode("classification") 

lgbm_wf <- 
  workflow() |> 
  add_model(lgbm_spec) |> 
  add_recipe(default_rec)

lgbm_grid <- grid_regular(learn_rate(), trees(), levels = c(4, 3))
``` 
 

## Fit models on training data

```{r}
# install.packages("kernlab")
library(kernlab)
```


```{r}
#| label: train

set.seed(9876)

Sys.time()

nn_res <- tune_grid(
  nn_wf,
  resamples = folds,
  grid = nn_grid,
  metrics = metric_set(roc_auc, accuracy, sens, spec),
   # save predictions
    control = control_grid(save_pred = TRUE)
)
nn_pred <- nn_res |>  collect_predictions()
 
Sys.time()


svm_res <- tune_grid(
  svm_wf,
  resamples = folds,
  grid = svm_grid,
  metrics = metric_set(roc_auc, accuracy, sens, spec),
   # save predictions
    control = control_grid(save_pred = TRUE)
)

svn_pred <- svm_res |>  collect_predictions()
 
Sys.time()


lgbm_res <- tune_grid(
  lgbm_wf,
  resamples = folds,
  grid = lgbm_grid,
  metrics = metric_set(roc_auc, accuracy, sens, spec),
   # save predictions
    control = control_grid(save_pred = TRUE)
)

lgbm_pred <- lgbm_res |>  collect_predictions()
 
Sys.time()

```



####  ROC curves for training


 
```{r}
#| label: roc-train-nn
#| out-width: "70%"

nn_pred |>
  group_by(id, hidden_units, penalty) |># id contains folds
  roc_curve(truth = default, .pred_Yes) |>
  autoplot() +
  facet_grid(rows = vars(hidden_units), cols = vars(penalty)) +
  theme(legend.position = "none")  +
  labs(title = "ROC Curve for Neural Network",
       subtitle = "Predicting Credit Card Default")
```

```{r}
#| label: roc-train-svm
#| out-width: "70%"

svn_pred |>
  group_by(id, cost, rbf_sigma) |># id contains folds
  roc_curve(truth = default, .pred_Yes) |>
  autoplot() +
  facet_grid(rows = vars(cost), cols = vars(rbf_sigma)) + 
  theme(legend.position = "none") +
  labs(title = "ROC Curve for Support Vector Machine",
       subtitle = "Predicting Credit Card Default")

```

```{r}
#| label: roc-train-lgbm
#| out-width: "70%"

lgbm_pred |>
  group_by(id, learn_rate, trees) |># id contains folds
  roc_curve(truth = default, .pred_Yes) |>
  autoplot() +
  facet_grid(rows = vars(learn_rate), cols = vars(trees)) +
  theme(legend.position = "none") +
  labs(title = "ROC Curve for LightGBM",
       subtitle = "Predicting Credit Card Default") 
```

**Plot metrics**

```{r}
#| label: train-metrics

nn_res |>
  collect_metrics() |>
  mutate(hidden_units = factor(hidden_units),
         penalty = factor(penalty)) |>
  ggplot(aes(hidden_units, mean, color = penalty, group = penalty)) +
  geom_line(linewidth = 1.5) +
  geom_point(size = 2) +
  scale_color_viridis_d(option = "plasma", begin = .9, end = 0)+
  facet_grid(rows =  vars(.metric))


svm_res |>
  collect_metrics() |>
  mutate(cost = factor(cost)
         , rbf_sigma = factor(rbf_sigma)
         ) |>
  ggplot(aes(cost, mean, color = rbf_sigma, group = rbf_sigma)) +
  geom_line(linewidth = 1.5) +
  geom_point(size = 2) +
  scale_color_viridis_d(option = "plasma", begin = .9, end = 0)+
  facet_grid(rows =  vars(.metric))

lgbm_res |>
  collect_metrics() |>
  mutate(learn_rate = factor(learn_rate)
         , trees = factor(trees)
         ) |>
  ggplot(aes(learn_rate, mean, color = trees, group = trees)) +
  geom_line(linewidth = 1.5) +
  geom_point(size = 2) +
  scale_color_viridis_d(option = "plasma", begin = .9, end = 0)+
  facet_grid(rows =  vars(.metric))
 
```
 
For the original parameter values: 

**NN**

* Penalty = 1 is consistently the best
* Hidden Units = 1 gives better accuracy and specificity, 4 gives better sensitivity



**SVM**

* rbf sigma = 1 is only value with sensible ROC
* Cost = lower two values are better than 32


**LightGBM**

* Trees = 1000 and 2000 give better ROC
* Learn Rate = 0.1 gives best ROC

## Select Final Model

```{r}
#| label: select-final


nn_best <- select_best(nn_res, metric = "roc_auc")

## alternatively, manually specify selected model
# nn_best <- tibble(
#   hidden_units = 1,
#   penalty  = 1,
#   rowNumber = 16, #obtain from nn_grid
#   .config = paste0("Preprocessor1_Model", 
#    ifelse(rowNumber<10, "0", ""), rowNumber)
# )
 

# Final NN model
nn_final_wf <- nn_wf |> 
  finalize_workflow(nn_best) 


# svm_best <- select_best(svm_res, metric = "roc_auc")

# Final SVM model
# svm_final_wf <- svm_wf |> finalize_workflow(svm_best) 


lgbm_best <- select_best(lgbm_res, metric = "roc_auc")

# Final SVM model
lgbm_final_wf <- lgbm_wf |> finalize_workflow(lgbm_best) 


```


# Evaluate
Do a final fit for each model (train on all training data and test on testing data)

```{r}
#| label: final-fit

nn_final_fit <- 
  nn_final_wf |>
  last_fit(split , metrics = metric_set(roc_auc, accuracy, sens, spec))


# svm_final_fit <- 
#   svm_final_wf |>
#   last_fit(split,   metrics = metric_set(roc_auc, accuracy, sens, spec))


lgbm_final_fit <- 
  lgbm_final_wf |>
  last_fit(split,   metrics = metric_set(roc_auc, accuracy, sens, spec))
```


## Assess performance
```{r}
#| label: collect-final-preds

nn_final_pred  <- nn_final_fit |> 
  collect_predictions()  |>
  mutate(model = "NN")

# svm_final_pred  <- svm_final_fit |> 
#   collect_predictions()   |>
#   mutate(model = "SVM")

lgbm_final_pred  <- lgbm_final_fit |> 
  collect_predictions()   |>
  mutate(model = "LGBM")


# Combine all predictions into a tibble
all_preds <- bind_rows(nn_final_pred, lgbm_final_pred)
```


Notice that there are only 7 defaults in the test dataset. 
Increasing the sample size is recommended in order to verify the model performance on a larger dataset.

### Metrics

```{r}
#| label: final-metrics

nn_final_fit |>   collect_metrics()
# svm_final_fit |>   collect_metrics()
lgbm_final_fit |>   collect_metrics()

```

When deciding which model is best, it is important to weigh up the 
importance of sensitivity vs specificity.
For example, is it worth it to correctly classify 2 extra customers as likely to "default",
if that means that an extra 50 customers will be incorrectly classified as likely to "default".

This trade-off will depend on the scenario and the costs involved in each situation.

### Confusion Matrix

```{r}
#| label: final-conf-mat

nn_final_fit$.predictions[[1]]  |>  
   conf_mat(truth = default, estimate = .pred_class)

# svm_final_fit$.predictions[[1]]  |>  
#    conf_mat(truth = default, estimate = .pred_class)

lgbm_final_fit$.predictions[[1]]  |>  
   conf_mat(truth = default, estimate = .pred_class)
```


### ROC for Test

```{r}
#| label: final-roc

all_preds |>
  group_by(model) |> 
  roc_curve(truth = default, .pred_Yes)  |> 
  autoplot() +
  ggtitle("ROC Curve: SVM,  NN vs LGBM")

```


# Alternative modelling decisions

A summary of results for alternative modelling decisions. Results may vary if different seeds are used.


* `step_downsample`, sample of 1000, 5 folds: 

   - computation time
     - NN: 5 sec
     - SVM: 5 sec
     - LightGBM: 15 sec
   - test performance (acc/sens/spec/roc_auc)
     - NN:  0.86, 0.714, 0.864, 0.893
     - SVM:  0.844, 0.714, 0.848, 0.896
     - LightGBM: 0.852, 0.571, 0.86, 0.812

* `step_upsample`, sample of 1000, 5 folds: 

   - computation time
     - NN: 15 sec
     - SVM: 15 sec
     - LightGBM: 45 sec
   - test performance (acc/sens/spec/roc_auc)
     - NN:  0.852, 0.714, 0.856, 0.889
     - SVM:  0.78, 0.714, 0.782, 0.895
     - LightGBM:  0.924, 0.571, 0.934, 0.757  


* `step_upsample`, whole dataset, 5 folds: 

   - approximate computation time
     - NN: 1 min
     - SVM: 8 min
     - LightGBM: 2.5 min 
   - test performance (acc/sens/spec/roc_auc)
     - NN:  0.870/0.819/0.872/0.940
     - SVM: 0.856/0.843/0.856/0.939
     - LightGBM: 0.963/0.361/0.983/0.892 (note poor sens)    

* `step_downsample`, whole dataset, 5 folds: 

   - approximate computation time
     - NN: 10 sec
     - SVM: 10 sec 
     - LightGBM: 45 sec
   - test performance (acc/sens/spec/roc_auc)
     - NN:   0.861, 0.819, 0.862, 0.938
     - SVM:  0.848, 0.831, 0.848, 0.931
     - LightGBM: 0.875, 0.831, 0.876, 0.905   
