---
format:
  pdf:
    geometry:
      - top=10mm
      - left=15mm
      - right=15mm
      - bottom=20mm
      - heightrounded
editor: visual
---

\begin{titlepage}
\vspace*{5cm}
\begin{center}
  {\LARGE\bfseries Understanding and Reducing Customer Churn at PlatefulNZ \par}
  \vspace{0.5em}
  {\large\itshape Data-Driven Insights for Improving Customer Retention in a Subscription-Based Meal Kit Service \par}
  \vspace{2em}
  {\large BUSINFO704 – Business Analytics \par}
  {\large Due: 19 August 2025 \par}
  \vspace{12em}
  {\large \textbf{Group Members:} \par}
  \vspace{0.5em}
  {\large Yicong Li – 691671170 \par}
  {\large Alice Wang – 123456789 \par}
  {\large John Zhang – 987654321 \par}
  {\large Emma Chen – 112233445 \par}
\end{center}
\end{titlepage}

```{r, echo=FALSE, results='hide', message=FALSE, warning=FALSE}
# ------------------------------
# install all packages
if (!require("tidyverse")) install.packages("tidyverse")
if (!require("knitr")) install.packages("knitr")
if (!require("tinytex")) install.packages("tinytex")
if (!require("tidymodels")) install.packages("tidymodels")
if (!require("skimr")) install.packages("skimr")
if (!require("ranger")) install.packages("ranger")
if (!require("car")) install.packages("car")
if (!require("vip")) install.packages("vip")
if (!require("bonsai")) install.packages("bonsai")     
if (!require("lightgbm")) install.packages("lightgbm") 
if (!require("xgboost")) install.packages("xgboost")
if (!require("broom")) install.packages("broom")
if (!require("klaR")) install.packages("klaR")
if (!require("discrim")) install.packages("discrim")
if (!require("patchwork")) install.packages("patchwork")
if (!require("yardstick")) install.packages("yardstick")
if (!require("glmnet")) install.packages("glmnet")
if (!require("FSelectorRcpp")) install.packages("FSelectorRcpp")
if (!require("themis")) install.packages("themis")
if (!require("psych")) install.packages("psych")


# ------------------------------
# library needs
library(tidymodels)
library(tidyverse)
library(knitr)
library(tinytex)
library(skimr)
library(psych)
library(vip)
library(bonsai)
library(ranger)
library(broom)
library(discrim)  
library(klaR)
library(patchwork)
library(yardstick)
library(glmnet)
library(FSelectorRcpp)
library(themis) 
library(car)
library(lightgbm)
library(xgboost)
                  

```

-- Introduction

-- Data Exploration & Methodology

```{r, echo=FALSE, results='hide', message=FALSE, warning=FALSE}
############ loading data & data processing ############
data <- read.csv("customers.csv")

# ---- Basic NA summary ----
missing_summary <- data %>%
  summarise(across(everything(),
                   ~ sum(is.na(.)),
                   .names = "na_{col}"))



# remove string 'week'
data$weeks_since_signup <- as.numeric(gsub(" weeks", "", data$weeks_since_signup))
data$weeks_since_last_purchase <- as.numeric(gsub(" weeks", "", data$weeks_since_last_purchase))


# remove string 'currency tag'
data$avg_AddOnpurchase_value <- as.numeric(gsub("[$]", "", data$avg_AddOnpurchase_value))




# gender type (3 types)
n_distinct(data$gender)
unique(data$gender) 
# "Female"   ///    "Prefer Not to Answer / Other / Unknown"    ///   "Male" 

data <- data %>% 
  mutate(
    gender = case_when(
      gender == "Female" ~ "Female",
      gender == "Male"   ~ "Male",
      TRUE               ~ "Other"  
    )
  )
unique(data$gender) 
# "Female"   ///   Other   ///   "Male" 

# subscription_payment_problem_last4Weeks (0 / 1)
data <- data %>%
  mutate(
    subscription_payment_problem_last4Weeks =
      as.integer(subscription_payment_problem_last4Weeks == "TRUE")
  )

# support ticket type (3 types)
n_distinct(data$support_ticket)
unique(data$support_ticket) 
# "NotLast3Months"  ////   "Last3Months"  ////    "YesThisMonth"


# sati survey type (6 types)
n_distinct(data$satisfaction_survey)
unique(data$satisfaction_survey) 
# 1 / 2 / 3 / 4 / 5 / NoResponse


# subscription type (5 types)
n_distinct(data$subscription)
unique(data$subscription) 
# "PlantBased" /// "Family"  //  "Everyday" ////  "Gourmet"  ////  "Low-carb"


# payment type (3 types)
n_distinct(data$payment)
unique(data$payment) 
# "DebitCard"  "CreditCard" "AppPay"  


#only take "weeks_since_last_purchase" equals or over 0
filter_data <- data %>% filter(weeks_since_last_purchase >= 0)


# View(data) haven't change
filter_data %>%
  count(retained_binary) %>%
  mutate(pct = n / sum(n) * 100)


# View(data) changed
filter_data <- filter_data %>%
  mutate(retained_binary = 1 - retained_binary)

filter_data %>%
  count(retained_binary) %>%
  mutate(pct = n / sum(n) * 100)
```

```{r, echo=FALSE, results='hide', message=FALSE, warning=FALSE}

categorical_vars <- c(
  "gender",
  "support_ticket",
  "opened_last_email",
  "subscription_payment_problem_last4Weeks",
  "last_login_device",
  "last_browser",
  "location",
  "subscription",
  "discounted_rate_last_purchase",
  "satisfaction_survey",
  "payment_type"
)

numeric_vars <- c(
  "app_visits",
  "website_visits",
  "social_media_comments",
  "social_media_posts",
  "social_media_likes",
  "social_media_shares",
  "subscription_meals_per_week",
  "subscription_people",
  "weeks_since_signup",
  "weeks_since_last_purchase",
  "num_purchases",
  "avg_AddOnpurchase_value"
)

```

```{r, echo=FALSE, results='hide', message=FALSE, warning=FALSE}

# ---- Numeric Variables Overview ----
num_summary <- psych::describe(filter_data %>% dplyr::select(dplyr::any_of(numeric_vars)))

# Output as table (knitr)
kable(num_summary, caption = "Summary statistics of numeric variables")

```

```{r, echo=FALSE, results='hide', message=FALSE, warning=FALSE}

# Normal distribution variables
nonSkewnessVariable <- c(
  "website_visits",
  "subscription_meals_per_week",
  "subscription_people",
  "weeks_since_signup",
  "num_purchases",
  "avg_AddOnpurchase_value"
)


# Skewed distribution variables
SkewnessVariables <- c(
  "app_visits",
  "social_media_comments",
  "social_media_posts",
  "social_media_likes",
  "social_media_shares",
  "weeks_since_last_purchase"
)
```

```{r, echo=FALSE, results='hide', message=FALSE, warning=FALSE}

# ---- Mann–Whitney Tests (Skewness test) ----
num_test_results <- map_dfr(SkewnessVariables, function(v) {
  ok <- complete.cases(filter_data[[v]], filter_data$retained_binary)
  group1 <- filter_data[[v]][ok & filter_data$retained_binary == 1]
  group0 <- filter_data[[v]][ok & filter_data$retained_binary == 0]
  
  test <- wilcox.test(group1, group0)
  
  tibble(
    variable = v,
    p_value = test$p.value,
    median_retained = median(group1, na.rm = TRUE),
    median_churned = median(group0, na.rm = TRUE)
  )
}) %>%
  arrange(p_value) %>%
  mutate(significant = ifelse(p_value < 0.05, "Yes", "No"))

# top 5 most significant numeric variables
head(num_test_results, 5)
```

```{r, echo=FALSE, results='hide', message=FALSE, warning=FALSE}

# ---- Homogeneity of variance test (Levene's Test) ----
levene_results <- lapply(nonSkewnessVariable, function(v) {
  formula <- as.formula(paste(v, "~ factor(retained_binary)"))
  test <- leveneTest(formula, data = filter_data)
  data.frame(
    variable = v,
    p_value = test$`Pr(>F)`[1],
    equal_variance = ifelse(test$`Pr(>F)`[1] > 0.05, "Yes", "No")
  )
})

levene_results <- do.call(rbind, levene_results)
levene_results
```

```{r, echo=FALSE, results='hide', message=FALSE, warning=FALSE}

# No significant mean difference
sameMeanVaribles <- c(
  "website_visits",
  "subscription_meals_per_week",
  "num_purchases"
)

# Significant mean difference
nonSameMeanVariables <- c(
  "subscription_people",
  "weeks_since_signup",
  "avg_AddOnpurchase_value"
)

```

```{r, echo=FALSE, results='hide', message=FALSE, warning=FALSE}


# sameMeanVariables → t-test
t_results_same <- lapply(sameMeanVaribles, function(v) {
  g1 <- filter_data[[v]][filter_data$retained_binary == 1]
  g0 <- filter_data[[v]][filter_data$retained_binary == 0]
  
  test <- t.test(g1, g0, var.equal = TRUE)
  
  data.frame(
    variable = v,
    method = "t-test (equal var)",
    p_value = test$p.value,
    mean_retained = mean(g1, na.rm = TRUE),
    mean_churned = mean(g0, na.rm = TRUE),
    mean_diff = mean(g1, na.rm = TRUE) - mean(g0, na.rm = TRUE)
  )
}) %>% bind_rows()


t_results_same
```

```{r, echo=FALSE, results='hide', message=FALSE, warning=FALSE}

# nonSameMeanVariables → Welch t-test
t_results_non_same <- lapply(nonSameMeanVariables, function(v) {
  g1 <- filter_data[[v]][filter_data$retained_binary == 1]
  g0 <- filter_data[[v]][filter_data$retained_binary == 0]
  
  test <- t.test(g1, g0, var.equal = FALSE)
  
  data.frame(
    variable = v,
    method = "Welch t-test",
    p_value = test$p.value,
    mean_retained = mean(g1, na.rm = TRUE),
    mean_churned = mean(g0, na.rm = TRUE),
    mean_diff = mean(g1, na.rm = TRUE) - mean(g0, na.rm = TRUE)
  )
}) %>% bind_rows()


t_results_non_same

```

```{r, echo=FALSE, results='hide', message=FALSE, warning=FALSE}

# gender - chi-test
tbl1 <- table(filter_data$gender, filter_data$retained_binary)
chisq.test(tbl1)

# support_ticket - chi-test
tbl2 <- table(filter_data$support_ticket, filter_data$retained_binary)
chisq.test(tbl2)

# open_last_email - chi-test
tbl3 <- table(filter_data$opened_last_email, filter_data$retained_binary)
chisq.test(tbl3)

# subscription_payment_problem_last4weeks - chi-test
tbl4 <- table(filter_data$subscription_payment_problem_last4Weeks, filter_data$retained_binary)
chisq.test(tbl4)

# last_login_device - chi-test
tbl5 <- table(filter_data$last_login_device, filter_data$retained_binary)
chisq.test(tbl5)

# last_browser - chi-test
tbl6 <- table(filter_data$last_browser, filter_data$retained_binary)
chisq.test(tbl6)

# location - chi-test
tbl7 <- table(filter_data$location, filter_data$retained_binary)
chisq.test(tbl7)

# subscription - chi-test
tbl8 <- table(filter_data$subscription, filter_data$retained_binary)
chisq.test(tbl8)

# payment_type - chi-test
tbl9 <- table(filter_data$payment_type, filter_data$retained_binary)
chisq.test(tbl9)

# discount_rate_last_purchase - chi-test
tbl10 <- table(filter_data$discounted_rate_last_purchase, filter_data$retained_binary)
chisq.test(tbl10)

```


```{r, echo=FALSE, results='hide', message=FALSE, warning=FALSE}

# gender - bar
genderBar <- ggplot(filter_data, aes(x = gender, fill = factor(retained_binary))) +
  geom_bar(position = "fill") +
  scale_y_continuous(labels = scales::percent) +
  scale_fill_manual(values = c("0" = "lightblue", "1" = "navyblue")) +
  labs(y = "Retention Rate", fill="if churn")
genderBar


# support_ticket - bar
supportTicketBar <- ggplot(filter_data, aes(x = support_ticket, fill = factor(retained_binary))) +
  geom_bar(position = "fill") +
  scale_y_continuous(labels = scales::percent) +
  scale_fill_manual(values = c("0" = "lightblue", "1" = "navyblue")) +
  labs(y = "Retention Rate", fill="if churn")
supportTicketBar

# opened_last_email - bar
openedLastEmailBar <- ggplot(filter_data, aes(x = opened_last_email, fill = factor(retained_binary))) +
  geom_bar(position = "fill") +
  scale_y_continuous(labels = scales::percent) +
  scale_fill_manual(values = c("0" = "lightblue", "1" = "navyblue")) +
  labs(y = "Retention Rate", fill="if churn")
openedLastEmailBar

# subscription_payment_problem_last4Weeks - bar
problemLast4WeeksBar <- ggplot(filter_data, aes(x = subscription_payment_problem_last4Weeks, fill = factor(retained_binary))) +
  geom_bar(position = "fill") +
  scale_y_continuous(labels = scales::percent) +
  scale_fill_manual(values = c("0" = "lightblue", "1" = "navyblue")) +
  labs(y = "Retention Rate", fill="if churn")
problemLast4WeeksBar

# last_login_device - bar
lastLoginDeviceBar <- ggplot(filter_data, aes(x = last_login_device, fill = factor(retained_binary))) +
  geom_bar(position = "fill") +
  scale_y_continuous(labels = scales::percent) +
  scale_fill_manual(values = c("0" = "lightblue", "1" = "navyblue")) +
  labs(y = "Retention Rate", fill="if churn")
lastLoginDeviceBar

# last_browser - bar
lastBrowserBar <- ggplot(filter_data, aes(x = last_browser, fill = factor(retained_binary))) +
  geom_bar(position = "fill") +
  scale_y_continuous(labels = scales::percent) +
  scale_fill_manual(values = c("0" = "lightblue", "1" = "navyblue")) +
  labs(y = "Retention Rate", fill="if churn")
lastBrowserBar

# location - bar
locationBar <- ggplot(filter_data, aes(x = location, fill = factor(retained_binary))) +
  geom_bar(position = "fill") +
  scale_y_continuous(labels = scales::percent) +
  scale_fill_manual(values = c("0" = "lightblue", "1" = "navyblue")) +
  labs(y = "Retention Rate", fill="if churn")
locationBar

# subscription - bar
subscriptionBar <- ggplot(filter_data, aes(x = subscription, fill = factor(retained_binary))) +
  geom_bar(position = "fill") +
  scale_y_continuous(labels = scales::percent) +
  scale_fill_manual(values = c("0" = "lightblue", "1" = "navyblue")) +
  labs(y = "Retention Rate", fill="if churn")
subscriptionBar

# payment_type - bar
paymentTypeBar <- ggplot(filter_data, aes(x = payment_type, fill = factor(retained_binary))) +
  geom_bar(position = "fill") +
  scale_y_continuous(labels = scales::percent) +
  scale_fill_manual(values = c("0" = "lightblue", "1" = "navyblue")) +
  labs(y = "Retention Rate", fill="if churn")
paymentTypeBar

# satisfaction_survey - bar
satisfactionSurveyBar <- ggplot(filter_data, aes(x = satisfaction_survey, fill = factor(retained_binary))) +
  geom_bar(position = "fill") +
  scale_y_continuous(labels = scales::percent) +
    scale_fill_manual(values = c("0" = "lightblue", "1" = "navyblue")) +
  labs(y = "Retention Rate", fill="if churn")
satisfactionSurveyBar

# discounted_rate_last_purchase - bar
discountedRateLastPurchaseBar <- ggplot(filter_data,aes(x = discounted_rate_last_purchase, fill = factor(retained_binary))) +
  geom_bar(position = "fill") +
  scale_y_continuous(labels = scales::percent) +
  scale_fill_manual(values = c("0" = "lightblue", "1" = "navyblue")) +
  labs(y = "Retention Rate", fill="if churn")
discountedRateLastPurchaseBar
```


```{r, echo=FALSE, results='hide', message=FALSE, warning=FALSE}

#final selective variables
finalVariables <- c(
  "weeks_since_last_purchase",
  "discounted_rate_last_purchase",
  "avg_AddOnpurchase_value",
  "num_purchases",
  "location",
  "last_browser",
  "subscription_payment_problem_last4Weeks",
  "satisfaction_survey"
)

```

```{r, echo=FALSE, results='hide', message=FALSE, warning=FALSE}

filter_data <- filter_data %>%
  mutate(
    retained_binary = factor(retained_binary, levels = c(0, 1))) %>% 
  dplyr::select(retained_binary, all_of(finalVariables))


set.seed(704)
# ---------- split data 80/20 ----------
split <- initial_split(filter_data, prop = 0.8, strata = retained_binary)
train_data <- training(split)
test_data  <- testing(split)

set.seed(704)
subsplit <- initial_split(train_data,
                          prop = 10000 / nrow(train_data),
                          strata = retained_binary)
train_sample <- training(subsplit)
nrow(train_sample)   # 10000 n as train sample

```

```{r, echo=FALSE, results='hide', message=FALSE, warning=FALSE}

# train smaple recipe
rec_tune <- recipe(retained_binary ~ ., data = train_sample) %>%
  step_string2factor(all_nominal_predictors()) %>%                # Convert string predictors to factors  
  step_unknown(all_nominal_predictors()) %>%                      # Handle unseen levels as "unknown" 
  step_other(all_nominal_predictors(), threshold = 0.01) %>%      # Group rare categories into "other"
  step_impute_median(all_numeric_predictors()) %>%                # Impute missing numeric values with median
  step_normalize(all_numeric_predictors())  %>%                   # Normalize numeric predictors (mean=0, sd=1)
  step_dummy(all_nominal_predictors()) %>%                        # Create dummy variables for categorical predictors
  step_zv(all_predictors()) %>%                                   # Remove predictors with zero variance
  step_upsample(retained_binary, over_ratio = 1, skip = TRUE)     # Balance classes by upsampling minority

set.seed(704)
# fold 5 times
folds <- vfold_cv(train_sample, v = 5, strata = retained_binary)

```


```{r, echo=FALSE, results='hide', message=FALSE, warning=FALSE}

# -------------------
# Training Function for LightGBM / Xboost / Random Forest / Naive Bayes
# -------------------

run_model <- function(model_spec, grid_spec, folds, train, test, model_name, recipe_spec) {
  wf <- workflow() %>% add_recipe(recipe_spec) %>% add_model(model_spec)

  set.seed(704)
  res <- tune_grid(
    wf, resamples = folds, grid = grid_spec,
    metrics = metric_set(roc_auc, accuracy, sens, spec, precision, f_meas), 
    control = control_grid(save_pred = TRUE, save_workflow = TRUE)
  )

  best     <- select_best(res, metric = "roc_auc")
  final_wf <- finalize_workflow(wf, best)
  fit_tr   <- fit(final_wf, data = train)
  preds_te <- augment(fit_tr, new_data = test)
  
  message("Training on: ", nrow(train), " rows | Testing on: ", nrow(test), " rows")

  test_metrics <- dplyr::bind_rows(
    roc_auc  (preds_te, truth = retained_binary, .pred_1,     event_level = "second"),    # note
    accuracy (preds_te, truth = retained_binary, .pred_class, event_level = "second"),    # note
    sens     (preds_te, truth = retained_binary, .pred_class, event_level = "second"),    # note
    spec     (preds_te, truth = retained_binary, .pred_class, event_level = "second"),    # note
    precision(preds_te, truth = retained_binary, .pred_class, event_level = "second"),    # note
    f_meas   (preds_te, truth = retained_binary, .pred_class, event_level = "second")     # note
  )

  # ← 5 time fold, mean / std_err
  cv_metrics <- res %>%
    collect_metrics() %>%
    dplyr::semi_join(best, by = ".config") %>%
    dplyr::select(.metric, mean, std_err)

  cv_roc <- collect_predictions(res, summarize = FALSE) %>%
    dplyr::semi_join(best, by = ".config") %>%
    dplyr::mutate(retained_binary = factor(retained_binary, levels = c("0","1"))) %>%
    dplyr::group_by(id) %>%
    yardstick::roc_curve(truth = retained_binary, .pred_1, event_level = "second") %>%     # note
    dplyr::ungroup() %>%
    dplyr::mutate(model = model_name)

  list(best_params = best, fit_tr = fit_tr, cv_metrics = cv_metrics, test_metrics = test_metrics, cv_roc = cv_roc)
}

```

```{r, echo=FALSE, results='hide', message=FALSE, warning=FALSE}
#| cache: true
# -------------------
# 1. LightGBM
# -------------------

start_time <- Sys.time()

lgbm_result <- run_model(
  boost_tree(
    learn_rate     = tune(),
    trees          = tune(),
    tree_depth     = tune(),
    min_n          = tune(),
    loss_reduction = tune()
  ) %>% set_engine(
        "lightgbm",
        deterministic = TRUE,  
        num_threads = 1,       
        seed = 704,            
        data_random_seed = 704,
        feature_fraction_seed = 704,
        bagging_seed = 704
      ) %>% set_mode("classification"),
  
  grid_random(
    learn_rate(range = c(-3, -0.5)),
    trees(range = c(400L, 1000L)),
    tree_depth(range = c(3L, 6L)),
    min_n(range = c(2L, 20L)),
    loss_reduction(range = c(-4, 1)),
    size = 10
  ),
  
  folds,
  train_sample,
  test_data,
  "LightGBM",
  rec_tune
)

end_time <- Sys.time()
runtime <- end_time - start_time
print(runtime)

lgbm_result


```



```{r, echo=FALSE, results='hide', message=FALSE, warning=FALSE}

# Extract the fitted LightGBM model from the workflow
lgb_model <- extract_fit_parsnip(lgbm_result$fit_tr)$fit

# Get feature importance
imp <- lgb.importance(lgb_model)

# Keep only features with Gain > 0
used_features <- imp$Feature[imp$Gain > 0]

# Plot the top 10 important features
lgb_model_vars <- lgb.plot.importance(imp, top_n = 10)
lgb_model_vars

```

```{r, echo=FALSE, results='hide', message=FALSE, warning=FALSE}
#| cache: true
# -------------------
# 2. XGBoost
# -------------------

start_time <- Sys.time()

xgb_result <- run_model(
  boost_tree(
    trees          = tune(),
    tree_depth     = tune(),
    learn_rate     = tune(),
    mtry           = tune(),
    min_n          = tune(),
    loss_reduction = tune(),
    sample_size    = tune()
  ) %>% set_engine(
  "xgboost",
  seed = 704,   
  nthread = 1   
) %>% set_mode("classification"),
  
  grid_random(
    trees(range = c(400L, 1000L)),
    tree_depth(range = c(3L, 6L)),
    learn_rate(range = c(-3, -0.5)),
    finalize(mtry(), train_sample),
    min_n(range = c(2L, 20L)),
    loss_reduction(range = c(-4, 1)),
    sample_prop(range = c(0.6, 1.0)),
    size = 10
  ),
  
  folds,
  train_sample,
  test_data,
  "XGBoost",
  rec_tune
)

end_time <- Sys.time()
runtime <- end_time - start_time
print(runtime)

xgb_result

```

```{r, echo=FALSE, results='hide', message=FALSE, warning=FALSE}

# Extract the fitted XGboost model from the workflow
xgb_model <- extract_fit_parsnip(xgb_result$fit_tr)$fit

# Get feature importance
imp_xgb <- xgb.importance(model = xgb_model)

# Keep only features with Gain > 0
used_features_xgb <- imp_xgb$Feature[imp_xgb$Gain > 0]


# Plot the top 10 important features
xgb_model_vars <- xgb.plot.importance(imp_xgb, top_n = 10)
xgb_model_vars
```


```{r, echo=FALSE, results='hide', message=FALSE, warning=FALSE}
#| cache: true
# -------------------
# 3. Random Forest
# -------------------

start_time <- Sys.time()

rf_result <- run_model(
  rand_forest(
    mtry  = tune(),
    trees = tune(),
    min_n = tune()
  ) %>% set_engine(
        "ranger",
        seed = 704,
        num.threads = 1,
        importance = "impurity"    
      ) %>% set_mode("classification"),
  
  grid_random(
    finalize(mtry(), train_sample),
    trees(range = c(500L, 1500L)),
    min_n(range = c(2L, 20L)),
    size = 10
  ),
  
  folds,
  train_sample,
  test_data,
  "Random Forest",
  rec_tune
)

end_time <- Sys.time()
runtime <- end_time - start_time
print(runtime)


rf_result

```


```{r, echo=FALSE, results='hide', message=FALSE, warning=FALSE}

# Extract the fitted Random Forest model from the workflow
rf_model <- extract_fit_parsnip(rf_result$fit_tr)$fit
# Get variable importance scores
imp_rf <- rf_model$variable.importance
# Convert importance scores into a tidy tibble
imp_rf_tbl <- enframe(imp_rf, name = "Feature", value = "Importance")

# Select and plot the top 10 most important features
rf_model_vars <- imp_rf_tbl %>%
  arrange(desc(Importance)) %>%
  slice_head(n = 10) %>%
  ggplot(aes(x = reorder(Feature, Importance), y = Importance)) +
  geom_col() +
  coord_flip() +
  labs(title = "Top 0 Features - Ranger Random Forest")

rf_model_vars
```


```{r, echo=FALSE, results='hide', message=FALSE, warning=FALSE}
#| cache: true
# -------------------
# 4. Naive Bayes
# -------------------


start_time <- Sys.time()

nb_result <- run_model(
  naive_Bayes(
    smoothness = tune(),
    Laplace    = tune()
  ) %>% set_engine("klaR") %>% set_mode("classification"),
  
  grid_random(
    smoothness(range = c(0, 2)),
    Laplace(range = c(0, 2)),
    size = 10
  ),
  
  folds,
  train_sample,
  test_data,
  "Naive Bayes",
  rec_tune
)

end_time <- Sys.time()
runtime <- end_time - start_time
print(runtime)


nb_result

```


```{r, echo=FALSE, results='hide', message=FALSE, warning=FALSE}

# Extract the recipe (preprocessing steps) from the fitted workflow
rec  <- extract_recipe(nb_result$fit_tr)
# # Apply the recipe to the training data to produce the processed dataset
proc <- bake(rec, new_data = train_sample)   


ig <- information_gain(retained_binary ~ ., data = proc)  # 列: attributes, importance
top_dummy <- ig[order(-ig$importance), ][1:10, ]             # top 10
top_dummy

```


```{r, echo=FALSE, results='hide', message=FALSE, warning=FALSE}


# -------------------------------
# Logistic Regression Function
# -------------------------------


fit_logit <- function(rec_tune, train, test, fold_cv = NULL,
                      event_level = "second") {        # ← set default positive class as 1 = churn
  
  # Define model specification & workflow
  spec <- logistic_reg(mode = "classification") %>% set_engine("glm")
  wf   <- workflow() %>% add_recipe(rec_tune) %>% add_model(spec)

  wf_fit <- fit(wf, data = train)

  # extract workflow -> model_fit -> engine(glm)
  mf <- workflows::extract_fit_parsnip(wf_fit)
  glm_obj <- parsnip::extract_fit_engine(mf)

  # Coefficients/intercept & Odds Ratio (95% CI)
  coef_table <- broom::tidy(glm_obj, conf.int = TRUE, exponentiate = FALSE)
  or_table   <- broom::tidy(glm_obj, conf.int = TRUE, exponentiate = TRUE)
  intercept  <- dplyr::filter(coef_table, term == "(Intercept)")

  # Predictions on test set
  preds <- augment(wf_fit, new_data = test)

  # Evaluation metrics (explicit package names to avoid conflicts)
  metr_set <- yardstick::metric_set(
    yardstick::roc_auc, yardstick::accuracy, yardstick::sens,
    yardstick::spec, yardstick::precision, yardstick::f_meas
  )
  metr_tbl <- metr_set(
    preds,
    truth   = retained_binary,
    .pred_1,                        # use .pred_1 (predicted churn probability)
    estimate = .pred_class,
    event_level = event_level
  ) %>% dplyr::arrange(.metric)

  # 7) Confusion matrix & ROC curve
  cm <- yardstick::conf_mat(preds, truth = retained_binary, estimate = .pred_class)
  roc_df <- yardstick::roc_curve(preds, retained_binary, .pred_1, event_level = event_level)  # ← reivse
  roc_plot <- autoplot(roc_df) + ggtitle("Logistic (glm) — Test ROC") + theme_minimal()

  # 8) Cross-validation
  cv_res <- NULL
  cv_roc <- NULL     
  if (!is.null(fold_cv)) {
    cv_res <- fit_resamples(
      wf, resamples = fold_cv,
      metrics = yardstick::metric_set(
        yardstick::roc_auc, yardstick::accuracy, yardstick::sens,
        yardstick::spec, yardstick::precision, yardstick::f_meas
      ),
      control = control_resamples(save_pred = TRUE)
    )
    
    
    # Collect predictions from each fold and compute ROC
    cv_preds <- collect_predictions(cv_res)
    cv_roc <- cv_preds %>%
      dplyr::group_by(id) %>%
      yardstick::roc_curve(retained_binary, .pred_1, event_level = event_level) %>%   # ← 修改5：CV 的 ROC 用 .pred_0
      dplyr::ungroup() %>%
      dplyr::mutate(model = "Logistic", id = as.factor(id))
  }

  # Return results
  list(
    wf_fit      = wf_fit,        # Trained workflow
    glm_obj     = glm_obj,       # Underlying glm object (use summary(glm_obj) to view intercept/coefficients)
    coef_table  = coef_table,    # Log-odds coefficients table (includes intercept)
    intercept   = intercept,     # Intercept row only
    odds_ratio  = or_table,      # Odds ratios & 95% CI
    metrics     = metr_tbl,      # Performance metrics
    conf_mat    = cm,            # Confusion matrix
    preds       = preds,         # Predictions on test set
    roc_curve_df= roc_df,        # ROC curve data
    cv_res      = cv_res,        # Cross-validation results (if any)
    cv_roc      = cv_roc         # Cross-validation ROC curve data
  )
}


```




```{r, echo=FALSE, results='hide', message=FALSE, warning=FALSE}
#| cache: true
# -------------------
# Logistic Regression
# -------------------

#  1w -- train sample

start_time <- Sys.time()

logit_result <- fit_logit(rec_tune, train_sample, test_data, fold_cv = folds)

end_time <- Sys.time(); 
print(end_time - start_time)

# Intercept, coefficients, and odds ratios
logit_result$intercept
logit_result$coef_table
logit_result$odds_ratio
# 
# Classic glm summary output (includes intercept)
# summary(logit_result$glm_obj)
# 
# Metrics, confusion matrix, and ROC curve
logit_result$metrics
logit_result$conf_mat
# 
# If CV was performed, check cross-validation metrics
collect_metrics(logit_result$cv_res)

```




```{r, echo=FALSE, results='hide', message=FALSE, warning=FALSE}

# -------------------
# Combine all ROC data
# -------------------
cv_roc_all <- bind_rows(
  lgbm_result$cv_roc,
  xgb_result$cv_roc,
  rf_result$cv_roc,
  nb_result$cv_roc,
  logit_result$cv_roc
)


```


```{r}

# Define a unified plotting function
plot_roc_single <- function(df, model_name) {
  ggplot(df, aes(x = 1 - specificity, y = sensitivity, color = id, group = id)) +
    geom_path(alpha = 0.5, linewidth = 0.5) +
    geom_abline(linetype = "dashed", color = "grey50") +
    labs(
      title = model_name ,
      x = "False Positive Rate (1 - Specificity)",
      y = "True Positive Rate (Sensitivity)"
    ) +
    coord_cartesian(xlim = c(0, 1), ylim = c(0, 1), expand = FALSE) +
    theme_minimal(base_size = 10) +
    theme(
      legend.position = "none",
      plot.title = element_text(face = "bold", hjust = 0.5, size = 11),
      axis.title.x = element_text(size = 9),  
      axis.title.y = element_text(size = 9),  
      axis.text = element_text(size = 8),     
      plot.margin = margin(6, 8, 6, 8)        
    )
}

# Generate one ROC plot per model
models <- unique(cv_roc_all$model)
p_list <- lapply(models, function(m) {
  plot_roc_single(dplyr::filter(cv_roc_all, model == m), m)
})

# Arrange plots into a grid (3 columns layout, automatic row wrapping)
fiveModels <- wrap_plots(p_list, ncol = 3)
fiveModels


```




```{r, echo=FALSE, results='hide', message=FALSE, warning=FALSE}

# best - logistic regression

# ---------- recipe + fold ----------
# 
rec_tune_final <- recipe(retained_binary ~ ., data = train_data) %>%
  step_string2factor(all_nominal_predictors()) %>%                  # Convert character vars → factors
  step_unknown(all_nominal_predictors()) %>%                        # Replace missing categorical levels with "unknown"
  step_other(all_nominal_predictors(), threshold = 0.01) %>%        # Group rare categories (<1%) into "other"
  step_impute_median(all_numeric_predictors()) %>%                  # Impute numeric predictors with median
  step_normalize(all_numeric_predictors())  %>%                     # Standardize numeric predictors (mean=0, sd=1)
  step_dummy(all_nominal_predictors()) %>%                          # One-hot encode categorical predictors
  step_zv(all_predictors()) %>%                                     # Remove zero-variance predictors
  step_downsample(retained_binary, under_ratio = 1, skip = TRUE)    # Downsample majority class

# --- Check original training set size & class distribution ---
nrow(train_data)
table(train_data$retained_binary)


# --- Prep recipe and apply downsampling ---
set.seed(704)
rec_prep  <- prep(rec_tune_final, training = train_data)
train_down <- juice(rec_prep)

# Check downsampled size & distribution
nrow(train_down)
table(train_down$retained_binary)

# fold 5 times
set.seed(704)
folds_final <- vfold_cv(train_data, v = 5, strata = retained_binary)

```


```{r, echo=FALSE, results='hide', message=FALSE, warning=FALSE}
#| cache: true
# -------------------
# Logistic Regression
# -------------------

#  16w -- train data

start_time <- Sys.time()

logit_result_final <- fit_logit(rec_tune_final, train_data, test_data, fold_cv = folds_final)

end_time <- Sys.time(); 
print(end_time - start_time)

# Intercept, coefficients, and odds ratios
logit_result_final$intercept
logit_result_final$coef_table
logit_result_final$odds_ratio


# Classic glm summary output (includes intercept)
# summary(logit_result$glm_obj)


# Metrics, confusion matrix, and ROC curve
logit_result_final$metrics
logit_result_final$conf_mat


# If CV was performed, check cross-validation metrics
collect_metrics(logit_result_final$cv_res)

```

```{r, echo=FALSE, results='hide', message=FALSE, warning=FALSE}


# Plot cross-validation ROC curves
p <- logit_result_final$cv_roc %>%
  dplyr::mutate(id = as.factor(id)) %>%
  dplyr::arrange(id, specificity) %>%
  ggplot(aes(x = 1 - specificity, y = sensitivity, color = id, group = id)) +
  geom_path(alpha = 0.5, linewidth = 0.5) +
  geom_abline(linetype = "dashed", color = "grey50") +
  labs(
    title = "Cross-Validation ROC Curves by Fold",
    x = "1 - Specificity (FPR)",
    y = "Sensitivity (TPR)"
  ) +
  theme_minimal(base_size = 14) +
  theme(legend.position = "none",
        strip.text = element_text(face = "bold", size = 14))

p


```

-- Findings & Result Analysis

-- Recommendations & Strategy

-- Conclusion

-- References

-- Appendices
