---
format:
  pdf:
    geometry:
      - top=10mm
      - left=15mm
      - right=15mm
      - bottom=20mm
      - heightrounded
editor: visual
---

\begin{titlepage}
\vspace*{5cm}
\begin{center}
  {\LARGE\bfseries Understanding and Reducing Customer Churn at PlatefulNZ \par}
  \vspace{0.5em}
  {\large\itshape Data-Driven Insights for Improving Customer Retention in a Subscription-Based Meal Kit Service \par}
  \vspace{2em}
  {\large BUSINFO704 – Business Analytics \par}
  {\large Due: 19 August 2025 \par}
  \vspace{12em}
  {\large \textbf{Group Members:} \par}
  \vspace{0.5em}
  {\large Yicong Li – 691671170 \par}
  {\large Alice Wang – 123456789 \par}
  {\large John Zhang – 987654321 \par}
  {\large Emma Chen – 112233445 \par}
\end{center}
\end{titlepage}

```{r, echo=FALSE, results='hide', message=FALSE, warning=FALSE}
if (!require("tidyverse")) install.packages("tidyverse")
if (!require("knitr")) install.packages("knitr")
if (!require("tinytex")) install.packages("tinytex")
if (!require("tidymodels")) install.packages("tidymodels")
if (!require("skimr")) install.packages("skimr")
library(tidymodels)
library(tidyverse)
library(knitr)
library(tinytex)
library(skimr)
library(psych)
```

-- Introduction

-- Data Exploration & Methodology

```{r, echo=FALSE, results='hide', message=FALSE, warning=FALSE}
############ loading data & data processing ############
data <- read.csv("customers.csv")

# ---- Basic NA summary ----
missing_summary <- data %>%
  summarise(across(everything(),
                   ~ sum(is.na(.)),
                   .names = "na_{col}"))



# remove string 'week'
data$weeks_since_signup <- as.numeric(gsub(" weeks", "", data$weeks_since_signup))
data$weeks_since_last_purchase <- as.numeric(gsub(" weeks", "", data$weeks_since_last_purchase))


# remove string 'currency tag'
data$avg_AddOnpurchase_value <- as.numeric(gsub("[$]", "", data$avg_AddOnpurchase_value))





# gender type (3 types)
n_distinct(data$gender)
unique(data$gender) 
# "Female"   ///    "Prefer Not to Answer / Other / Unknown"    ///   "Male" 

data <- data %>% 
  mutate(
    gender = case_when(
      gender == "Female" ~ "Female",
      gender == "Male"   ~ "Male",
      TRUE               ~ "Other"   # 把剩下的都归到 Other
    )
  )
unique(data$gender) 
# "Female"   ///   Other   ///   "Male" 


# support ticket type (3 types)
n_distinct(data$support_ticket)
unique(data$support_ticket) 
# "NotLast3Months"  ////   "Last3Months"  ////    "YesThisMonth"


# sati survey type (6 types)
n_distinct(data$satisfaction_survey)
unique(data$satisfaction_survey) 
# 1 / 2 / 3 / 4 / 5 / NoResponse

data <- data %>% 
  mutate(
    # 先把 NoResponse 改成字符 “0”
    satisfaction_survey = recode(
      satisfaction_survey,
      "NoResponse" = "0",
      .default     = satisfaction_survey
    ),
    # 再把整个列转成整数型
    satisfaction_survey = as.integer(satisfaction_survey)
  )
unique(data$satisfaction_survey) 
# 1 / 2 / 3 / 4 / 5 / NoResponse (0)


# subscription type (5 types)
n_distinct(data$subscription)
unique(data$subscription) 
# "PlantBased" /// "Family"  //  "Everyday" ////  "Gourmet"  ////  "Low-carb"


# payment type (3 types)
n_distinct(data$payment)
unique(data$payment) 
# "DebitCard"  "CreditCard" "AppPay"  


#only take "weeks_since_last_purchase" equals or over 0
filter_data <- data %>% filter(weeks_since_last_purchase >= 0)


# View(data)
filter_data %>%
  count(retained_binary) %>%
  mutate(pct = n / sum(n) * 100)
```

```{r, echo=FALSE, results='hide', message=FALSE, warning=FALSE}

# 
categorical_vars <- c(
  "gender",
  "support_ticket",
  "opened_last_email",
  "subscription_payment_problem_last4Weeks",
  "last_login_device",
  "last_browser",
  "location",
  "subscription",
  "payment_type"
)

numeric_vars <- c(
  "app_visits",
  "website_visits",
  "social_media_comments",
  "social_media_posts",
  "social_media_likes",
  "social_media_shares",
  "subscription_meals_per_week",
  "subscription_people",
  "weeks_since_signup",
  "weeks_since_last_purchase",
  "num_purchases",
  "discounted_rate_last_purchase",
  "avg_AddOnpurchase_value"
)

```


```{r, echo=FALSE, results='hide', message=FALSE, warning=FALSE}

# ---- 数值型概览 ----
num_summary <- psych::describe(filter_data %>% select(any_of(numeric_vars)))

# 输出成表（knitr）
kable(num_summary, caption = "数值变量描述统计")

```

```{r, echo=FALSE, results='hide', message=FALSE, warning=FALSE}

# 批量 Mann–Whitney 检验
num_test_results <- map_dfr(numeric_vars, function(v) {
  ok <- complete.cases(filter_data[[v]], filter_data$retained_binary)
  group1 <- filter_data[[v]][ok & filter_data$retained_binary == 1]
  group0 <- filter_data[[v]][ok & filter_data$retained_binary == 0]
  
  test <- wilcox.test(group1, group0)
  
  tibble(
    variable = v,
    p_value = test$p.value,
    median_retained = median(group1, na.rm = TRUE),
    median_churned = median(group0, na.rm = TRUE)
  )
}) %>%
  arrange(p_value) %>%
  mutate(significant = ifelse(p_value < 0.05, "Yes", "No"))

# 看前 5 个最显著的数值变量
head(num_test_results, 5)
```

```{r}

library(ggplot2)

# 定义阈值（这里用中位数）
x_median <- median(filter_data$weeks_since_last_purchase, na.rm = TRUE)
y_median <- median(filter_data$num_purchases, na.rm = TRUE)

ggplot(filter_data, aes(
  x = weeks_since_last_purchase,
  y = num_purchases,
  color = factor(retained_binary)
)) +
  geom_point(alpha = 0.4) +
  geom_vline(xintercept = x_median, linetype = "dashed", color = "grey40") +
  geom_hline(yintercept = y_median, linetype = "dashed", color = "grey40") +
  annotate("text", x = x_median * 0.5, y = y_median * 1.5, label = "高购买 / 近期活跃", color = "black") +
  annotate("text", x = x_median * 1.5, y = y_median * 1.5, label = "高购买 / 久未购买", color = "black") +
  annotate("text", x = x_median * 0.5, y = y_median * 0.5, label = "低购买 / 近期活跃", color = "black") +
  annotate("text", x = x_median * 1.5, y = y_median * 0.5, label = "低购买 / 久未购买", color = "black") +
  scale_color_manual(values = c("0" = "red", "1" = "blue"),
                     labels = c("流失", "留存")) +
  labs(
    x = "距离上次购买的周数",
    y = "总购买次数",
    color = "留存状态",
    title = "客户分布象限图",
    subtitle = "按总购买次数和最近购买时间划分客户类型"
  ) +
  theme_minimal()

```


```{r, echo=FALSE, results='hide', message=FALSE, warning=FALSE}

# 以 gender 为例
tbl1 <- table(filter_data$gender, filter_data$retained_binary)
chisq.test(tbl1)

# 以 support_ticket 为例
tbl2 <- table(filter_data$support_ticket, filter_data$retained_binary)
chisq.test(tbl2)

# 以 open_last_email 为例
tbl3 <- table(filter_data$opened_last_email, filter_data$retained_binary)
chisq.test(tbl3)

# 以 subscription_payment_problem_last4weeks 为例
tbl4 <- table(filter_data$subscription_payment_problem_last4Weeks, filter_data$retained_binary)
chisq.test(tbl4)

# 以 last_login_device 为例
tbl5 <- table(filter_data$last_login_device, filter_data$retained_binary)
chisq.test(tbl5)

# 以 last_browser 为例
tbl6 <- table(filter_data$last_browser, filter_data$retained_binary)
chisq.test(tbl6)

# 以 location 为例
tbl7 <- table(filter_data$location, filter_data$retained_binary)
chisq.test(tbl7)

# 以 subscription 为例
tbl8 <- table(filter_data$subscription, filter_data$retained_binary)
chisq.test(tbl8)

# 以 payment_type 为例
tbl9 <- table(filter_data$payment_type, filter_data$retained_binary)
chisq.test(tbl9)

```

```{r, echo=FALSE, results='hide', message=FALSE, warning=FALSE}

#
ggplot(filter_data, aes(x = gender, fill = factor(retained_binary))) +
  geom_bar(position = "fill") +
  scale_y_continuous(labels = scales::percent) +
  labs(y = "Retention Rate", fill = "Retained (1=Yes, 0=No)")

#
ggplot(filter_data, aes(x = support_ticket, fill = factor(retained_binary))) +
  geom_bar(position = "fill") +
  scale_y_continuous(labels = scales::percent) +
  labs(y = "Retention Rate", fill = "Retained (1=Yes, 0=No)")

#
ggplot(filter_data, aes(x = opened_last_email, fill = factor(retained_binary))) +
  geom_bar(position = "fill") +
  scale_y_continuous(labels = scales::percent) +
  labs(y = "Retention Rate", fill = "Retained (1=Yes, 0=No)")


#
ggplot(filter_data, aes(x = subscription_payment_problem_last4Weeks, fill = factor(retained_binary))) +
  geom_bar(position = "fill") +
  scale_y_continuous(labels = scales::percent) +
  labs(y = "Retention Rate", fill = "Retained (1=Yes, 0=No)")


#
ggplot(filter_data, aes(x = last_login_device, fill = factor(retained_binary))) +
  geom_bar(position = "fill") +
  scale_y_continuous(labels = scales::percent) +
  labs(y = "Retention Rate", fill = "Retained (1=Yes, 0=No)")


#
ggplot(filter_data, aes(x = last_browser, fill = factor(retained_binary))) +
  geom_bar(position = "fill") +
  scale_y_continuous(labels = scales::percent) +
  labs(y = "Retention Rate", fill = "Retained (1=Yes, 0=No)")

#
ggplot(filter_data, aes(x = location, fill = factor(retained_binary))) +
  geom_bar(position = "fill") +
  scale_y_continuous(labels = scales::percent) +
  labs(y = "Retention Rate", fill = "Retained (1=Yes, 0=No)")


#
ggplot(filter_data, aes(x = subscription, fill = factor(retained_binary))) +
  geom_bar(position = "fill") +
  scale_y_continuous(labels = scales::percent) +
  labs(y = "Retention Rate", fill = "Retained (1=Yes, 0=No)")


#
ggplot(filter_data, aes(x = payment_type, fill = factor(retained_binary))) +
  geom_bar(position = "fill") +
  scale_y_continuous(labels = scales::percent) +
  labs(y = "Retention Rate", fill = "Retained (1=Yes, 0=No)")
```

-- Findings & Result Analysis

-- Recommendations & Strategy

-- Conclusion

-- References

-- Appendices
