---
format:
  pdf:
    geometry:
      - top=10mm
      - left=15mm
      - right=15mm
      - bottom=20mm
      - heightrounded
editor: visual
---

\begin{titlepage}
\vspace*{5cm}
\begin{center}
  {\LARGE\bfseries Understanding and Reducing Customer Churn at PlatefulNZ \par}
  \vspace{0.5em}
  {\large\itshape Data-Driven Insights for Improving Customer Retention in a Subscription-Based Meal Kit Service \par}
  \vspace{2em}
  {\large BUSINFO704 – Business Analytics \par}
  {\large Due: 19 August 2025 \par}
  \vspace{12em}
  {\large \textbf{Group Members:} \par}
  \vspace{0.5em}
  {\large Yicong Li – 691671170 \par}
  {\large Alice Wang – 123456789 \par}
  {\large John Zhang – 987654321 \par}
  {\large Emma Chen – 112233445 \par}
\end{center}
\end{titlepage}

```{r, echo=FALSE, results='hide', message=FALSE, warning=FALSE}
if (!require("tidyverse")) install.packages("tidyverse")
if (!require("knitr")) install.packages("knitr")
if (!require("tinytex")) install.packages("tinytex")
if (!require("tidymodels")) install.packages("tidymodels")
if (!require("skimr")) install.packages("skimr")
library(tidymodels)
library(tidyverse)
library(knitr)
library(tinytex)
library(skimr)
```

-- Introduction

-- Data Exploration & Methodology

```{r, echo=FALSE, results='hide', message=FALSE, warning=FALSE}
############ loading data & data processing ############
data <- read.csv("customers.csv")
# View(data)
# skim(customerData)


# remove string 'week'
data$weeks_since_signup <- as.numeric(gsub(" weeks", "", data$weeks_since_signup))
data$weeks_since_last_purchase <- as.numeric(gsub(" weeks", "", data$weeks_since_last_purchase))


# remove string 'currency tag'
data$avg_AddOnpurchase_value <- as.numeric(gsub("[$]", "", data$avg_AddOnpurchase_value))


# gender type (3 types)
n_distinct(data$gender)
unique(data$gender) 
# "Female"   ///    "Prefer Not to Answer / Other / Unknown"    ///   "Male" 


# support ticket type (3 types)
n_distinct(data$support_ticket)
unique(data$support_ticket) 
# "NotLast3Months"  ////   "Last3Months"  ////    "YesThisMonth"


# sati survey type (6 types)
n_distinct(data$satisfaction_survey)
unique(data$satisfaction_survey) 
# 1 / 2 / 3 / 4 / 5 / NoResponse


# subscription type (5 types)
n_distinct(data$subscription)
unique(data$subscription) 
# "PlantBased" /// "Family"  //  "Everyday" ////  "Gourmet"  ////  "Low-carb"


# payment type (3 types)
n_distinct(data$payment)
unique(data$payment) 
# "DebitCard"  "CreditCard" "AppPay"  
```

-- Findings & Result Analysis

-- Recommendations & Strategy

-- Conclusion

-- References

-- Appendices
