---
format:
  pdf:
    geometry:
      - top=10mm
      - left=15mm
      - right=15mm
      - bottom=20mm
      - heightrounded
editor: visual
---

--Cover Page

\begin{titlepage}
\vspace*{5cm}
\begin{center}
  {\LARGE\bfseries Understanding and Reducing Customer Churn at PlatefulNZ \par}
  \vspace{0.5em}
  {\large\itshape Data-Driven Insights for Improving Customer Retention in a Subscription-Based Meal Kit Service \par}
  \vspace{2em}
  {\large BUSINFO704 – Business Analytics \par}
  {\large Due: 19 August 2025 \par}
  \vspace{12em}
  {\large \textbf{Group Members:} \par}
  \vspace{0.5em}
  {\large Yicong Li – 691671170 \par}
  {\large Alice Wang – 123456789 \par}
  {\large John Zhang – 987654321 \par}
  {\large Emma Chen – 112233445 \par}
\end{center}
\end{titlepage}

```{r, echo=FALSE, results='hide', message=FALSE, warning=FALSE}
if (!require("tidyverse")) install.packages("tidyverse")
if (!require("knitr")) install.packages("knitr")
if (!require("tinytex")) install.packages("tinytex")
if (!require("tidymodels")) install.packages("tidymodels")
library(tidymodels)
library(tidyverse)
library(knitr)
library(tinytex)
```

-- Introduction

-- Data Exploration & Methodology

```{r, echo=FALSE, results='hide', message=FALSE, warning=FALSE}
customerMetaData <- read.csv("customers_metadata.csv")
customerData <- read.csv("customers.csv")
View(customerMetaData)
View(customerData)
```



-- Findings & Result Analysis

-- Recommendations & Strategy

-- Conclusion

-- References

-- Appendices
